A sample computation

### 2024-05-26 draft_systoles.tex
test update draft_systoles.tex
add bounds, family, the unknown

\DeclareMathOperator{\assign}{\coloneqq}        % Assingment operator
\newcommand{\R}{\mathbb{R}}        % Real
\newcommand{\Q}{\mathbb{Q}}        % rationals
\newcommand{\C}{\mathbb{C}}        % Complex
\renewcommand{\P}{\mathbb{P}}        % Projective
\newcommand{\A}{\mathbb{A}}        % Affine
\renewcommand{\O}{\mathcal{O}}        % Structure sheaf
\newcommand{\Id}{\mathbf{Id}}        % Identity
\newcommand{\SL}{\mathbf{SL}_3(\mathbb{Z})}        % SL_3(Z)
\newcommand{\Mat}{\mathbf{Mat}_3(\mathbb{C})}        % Mat_3( C )
\newcommand{\SLp}{\Gamma_p}        % a congruence subgroup
\newcommand{\phip}{\phi_p}        % the golden ratio mod(p)

@@

@@

\section{Intro}
Recall Lucas primes (see \url{https://en.wikipedia.org/wiki/Lucas_number#Lucas_primes}, \url{https://t5k.org/top20/page.php?id=48})
\begin{equation}
	\label{the_primes_sequence}
2, 3, 7, 11, 29, 47, 199, 521, 2207, 3571, 9349, 3010349, 54018521, 370248451, 6643838879, \ldots
\end{equation}
In this notes we establish the following results:
\begin{enumerate}
\item
The regularor of $\tilde A$
grows as $\approx \ln^2(p)$.

@@

Notice that 
\begin{multline}
\chi_{\tilde A}(\lambda) 
= \det(\lambda \Id - \tilde A) \\
= \det(\lambda \Id - (\Id + p A))
= \det((\lambda - 1) \Id - p A)\\
= p^3 \det(\frac{\lambda - 1}p \Id - A)
= p^3 \chi_A(\frac{\lambda - 1}p)
.\end{multline}
Hence adding the root of $\chi_{A}$ or $\chi_{\tilde A}$ result in the same field; therefore we reduce to showing that 
$\Reg ( \Q(\alpha)/\Q ) \approx \ln^2(p) $ for $\alpha$ a root of $\chi_{A}(t) = f(t) = t^3 - (2p t^2 + (p + 2) t + 1) $.

@@

@@

>>>>

\DeclareMathOperator{\assign}{\coloneqq}        % Assingment operator
\newcommand{\R}{\mathbb{R}}        % Real
\newcommand{\Q}{\mathbb{Q}}        % rationals
\newcommand{\C}{\mathbb{C}}        % Complex
\renewcommand{\P}{\mathbb{P}}        % Projective
\newcommand{\A}{\mathbb{A}}        % Affine
\renewcommand{\O}{\mathcal{O}}        % Structure sheaf
\newcommand{\Id}{\mathbf{Id}}        % Identity
\newcommand{\SL}{\mathbf{SL}_3(\mathbb{Z})}        % SL_3(Z)
\newcommand{\GL}{\mathbf{GL}_3} % GL_3
\newcommand{\Mat}{\mathbf{Mat}_3(\mathbb{C})}        % Mat_3( C )
\newcommand{\MatZ}{\mathbf{Mat}_3(\mathbb{Z})}        % Mat_3( ZZ )
\newcommand{\SLp}{\Gamma_p}        % a congruence subgroup
\newcommand{\phip}{\phi_p}        % the golden ratio mod(p)
\newcommand{\hence}{\Rightarrow}        % hence

@@

\tableofcontents

@@

\section{Intro}
Recall the Lucas primes (see \url{https://en.wikipedia.org/wiki/Lucas_number#Lucas_primes}, \url{https://t5k.org/top20/page.php?id=48})
\begin{equation}
	\label{the_primes_sequence}
2, 3, 7, 11, 29, 47, 199, 521, 2207, 3571, 9349, 3010349, 54018521, 370248451, 6643838879, \ldots
\end{equation}
In this notes we establish the following results:
\begin{enumerate}
\item
The regularor of $\tilde A$
grows as $\approx \ln^2(p)$.

@@

Notice that 
\begin{multline}
	\label{computation_char_poly_of_a_tilde}
\chi_{\tilde A}(\lambda) 
= \det(\lambda \Id - \tilde A) \\
= \det(\lambda \Id - (\Id + p A))
= \det((\lambda - 1) \Id - p A)\\
= p^3 \det(\frac{\lambda - 1}p \Id - A)
= p^3 \chi_A(\frac{\lambda - 1}p)
.\end{multline}
Hence adding the root of $\chi_{A}$ or $\chi_{\tilde A}$ result in the same field; therefore we reduce to showing that 
$\Reg ( \Q(\alpha)/\Q ) \approx \ln^2(p) $ for $\alpha$ a root of $\chi_{A}(t) = f(t) = t^3 - (2p t^2 + (p + 2) t + 1) $.

@@

\subsection{Lucas primes give small degree $\log(p)$}

@@

\subsection{Bounds on the degree $log(p) \prec \deg \prec p$}
Motivated by geometric considerations, we call the degree $\deg$ the index $[Z_{\SLp}(\tilde A) : Z_{\SL}(\tilde A)]$.

\begin{Prop}
We have $O(log(p)) \le \deg \le O(p)$.
\end{Prop}

\begin{proof}
The proof of the previous Theorem shows that $Z_{\SL}(\tilde A)$ is generated by  $\alpha^3(2 \alpha + 1)^{-1}, \alpha^{\deg}$.

So, the goal is to show that the multiplicative order of $\phi_p$ is bounded by $\log(p)$ from below and by $p$ from above.

First let us show $\log(p) < \deg$.
Indeed, lifting $\phi_p$ to $\phi$ we notice that 
\[
\phi_p^k = 1 \hence (- \phi_p^{-1} )^{k} = (-1)^k \hence \phi^k + (- \phi^{-1} )^{k} - (1 + (-1)^k) \in p \mathbb{Z}
.\] 
Thus the degree $\deg > \log(p) - \epsilon$ for $p$ big enough and some small $\epsilon$.

Finally, we shall show $\deg \le O(p)$.
What we show in reality is that if $\sqrt 5 \in \mathbb F_p$, the degree $\deg$ divides $p-1$, $\deg | p-1$ and if $\sqrt 5 \notin \mathbb F_p$, $\deg | 2 p + 2$.

Indeed, the first part is clear by Fermat's little theorem.
To show the second, let's decompose 
\[
x^2 - x - 1 = (x - \alpha_1)(x - \alpha_2)
,\] 
where $\alpha_1, \alpha_2 \in \mathbb F_{p^2}$.
The Frobenius automorphism $\mathrm{Frob}: t \mapsto t^p$ acts on the roots $\alpha_1, \alpha_2$ by permuting them; thus we have $\alpha_1^p = \alpha_2 = -1/\alpha_1$ implying $\alpha_1^{2p+2} = (-1)^2 = 1$.
\end{proof}

\section{Enter family}
Let $f_{a,b}(t) = t^3 - a p t^2 - (b p + a) t - b$.
Here $a, b \in \mathbb{Z} $.
Note that $f_{2,1}(t) = f(t)$ is the characteristic polynomial of $A$.

We can now define
\begin{equation}
\label{matrix_small_family}
A_{a, b} = 
\begin{pmatrix}
0 & 0 & b \\
1 & 0 & bp+a \\
0 & 1 & ap
\end{pmatrix}
\in \MatZ,
\end{equation}
\begin{equation}
\label{matrix_big_family}
\tilde A_{a, b} = \Id + p A_{a, b} =
\begin{pmatrix}
1 & 0 & b p \\
p & 1 & b p^2 + a p \\
0 & p & 1 + a p^2
\end{pmatrix}
\in \SLp.
\end{equation}
Here $\tilde A_{a, b}$ is indeed invertible thanks to 
\begin{multline}
	\label{computation_char_poly_of_a_tilde_family}
	\det({{\tilde{A}}_{a,b}})
= p^3 \det \left(\frac{1}p \Id + A_{a,b}\right) 
= - p^3 \det \left(\frac{-1}p \Id - A_{a,b}\right) \\
= - p^3 f_{a,b}\left(\frac{-1}p\right)
= 1 + a p^2 - (b p + a) p^2 + b p^3
= 1
.\end{multline}
Actually, almost all of the matrices from $\SLp$ are similar to $\tilde A_{p, q}$.
\begin{Prop}
Any matrix $B \in \SLp$ whose minimal polynomial is its characteristic polynomial satisfies
\[
B = C \tilde A_{a, b} C^{-1} 
,\] for some $a, b \in \mathbb{Z}, C \in \GL( \mathbb{Q} )$.
\end{Prop}
\begin{proof}
Pick $C \in \GL( \mathbb{Q} ) $ such that 
$$
C^{-1} \frac{1}{p}(B - \Id) C =
\begin{pmatrix}
0 & 0 & r_0 \\
1 & 0 & r_1 \\
0 & 1 & r_2
\end{pmatrix} \in 
\mathbf{SL}_3(\mathbb{Q})
.$$ 
This is possible to do because of the condition on $B$ and the theory of Frobenius normal form over $ \mathbb{Q}$ from linear algebra.

This matrix is actually integer $ C^{-1} \frac{1}{p}(B - \Id) C \in \MatZ$.
Indeed, the coefficients in the last column are the coefficients of the characteristic polynomial. 
Therefore, $ C^{-1} B C \in \SLp $.

It is left to show that 
$ C^{-1} B C = \tilde A_{a, b}$.
This follows from the equation $\det C^{-1} B C = 1$ which reads
\[
1 + r_2 p - r_1 p^2 + r_0 p^3 = 1
,\] 
which implies
 \[
r_0 = b, r_1 = bp + a, r_2 = ap,
\] 
for some $a, b \in \mathbb{Z}$.
\end{proof}

\section{The Unknown}
\begin{Prob}
What is the behavior of $\Reg, \deg $ for an arbitrary member of the family $\tilde A_{a, b}$?
\end{Prob}
Currently I don't really understand how to treat ramification.
